---
title: "Auditing Feature Matches"
description: "Inspect exact and sentence-level feature evidence with audit_features() using pre-parsed UDPipe data."
page-layout: article
format:
  html:
    toc: true
engine: knitr
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
library(pseudobibeR.fr)
library(udpipe)
library(dplyr)
library(stringr)

load("../data/french_corpus.rda")   # loads 'french_corpus'
load("../data/fc_udpipe_parsed.rda")   # loads 'fc_udpipe_parsed'

discipline_lookup <- french_corpus |>
  distinct(doc_id, discipline)
```

## Why audit feature matches?

`biber()` gives you document-level counts. `audit_features()` lets you step down
to the sentence or token span that triggered those counts so you can review
edge cases, confirm parser behavior, and share examples with collaborators.

The audit API now supports three common review modes:

- Inspect a single feature with exact KWIC-style anchors when available
- Sample a whole feature category without reading every sentence in the corpus
- Distinguish between exact token-span matches and sentence-level fallback rows

::: {.callout-note}
## Data
This vignette uses the Chambers-Le Baron UDPipe parse stored in
`_quarto/data/fc_udpipe_parsed.rda`, plus the companion document metadata in
`_quarto/data/french_corpus.rda`.
:::

## Corpus snapshot

```{r corpus-summary}
tibble::tibble(
  documents = nrow(french_corpus),
  disciplines = n_distinct(french_corpus$discipline),
  parsed_tokens = nrow(fc_udpipe_parsed)
)
```

## Start with one exact feature

We'll start with by-passives.

```{r passive-voice}
passive_audit <- audit_features(
  fc_udpipe_parsed,
  feature = "f_18_by_passives",
  sample_n = 8,
  max_per_doc = 1,
  seed = 1,
  window = 6
)

passive_audit
```

```{r particle-display}
passive_audit |>
  left_join(discipline_lookup, by = "doc_id") |>
  select(discipline, doc_id, sentence_id, keyword, left, right, chunk)
```

The compact print method gives you a quick summary of how many returned rows are
exact anchors versus sentence-level fallback rows.

## Audit a whole category

The category interface is useful when you want representative evidence across a
family of related features without choosing them one by one.

```{r subordination-category}
subordination_audit <- audit_features(
  fc_udpipe_parsed,
  category = "Subordination features",
  sample_n = 12,
  max_per_doc = 2,
  seed = 7,
  window = 5
)

subordination_audit |>
  left_join(discipline_lookup, by = "doc_id") |>
  select(discipline, feature, match_type, keyword, chunk)
```

```{r subordination-summary}
subordination_audit |>
  count(feature, match_type, sort = TRUE)
```

Because the audit locators now cover complement clauses, relatives, adverbial
subordinators, and reduced clause types, many rows in this category should be
returned with exact anchors.

## Compare exact and sentence-level review modes

Not every feature family has an exact locator yet. For those features,
`audit_features()` still returns sentence chunks so you can review what was
counted.

```{r sentence-fallback}
verb_class_audit <- audit_features(
  fc_udpipe_parsed,
  category = "Specialized verb classes",
  sample_n = 8,
  max_per_doc = 1,
  seed = 11
)

verb_class_audit |>
  select(feature, doc_id, sentence_id, match_type, keyword, chunk)
```

In this case, `match_type` helps you distinguish rows backed by exact token-span
 locators from rows returned through sentence-level fallback.

## Review reduced forms and negation together

You can also pass a vector of specific features when you want a targeted review
set for debugging or annotation meetings.

```{r targeted-review}
targeted_audit <- audit_features(
  fc_udpipe_parsed,
  feature = c(
    "f_60_that_deletion",
    "f_61_stranded_preposition",
    "f_62_split_infinitive",
    "f_63_split_auxiliary",
    "f_66_neg_synthetic",
    "f_67_neg_analytic"
  ),
  sample_n = 14,
  max_per_doc = 2,
  seed = 21,
  window = 7
)

targeted_audit |>
  left_join(discipline_lookup, by = "doc_id") |>
  select(discipline, feature, keyword, left, right, chunk)
```

This pattern is especially useful when you want to inspect recently added or
parser-sensitive heuristics side by side.

## Practical review tips

- Use `sample_n` when a feature is common enough that full inspection would be noisy
- Use `max_per_doc` to avoid letting one document dominate the sample
- Increase `window` for multiword constructions such as negation or split auxiliaries
- Join audit output back to corpus metadata so reviewers can stratify by discipline, register, or source
- Reserve `biber()` for quantitative summaries and `audit_features()` for qualitative spot-checking

## Next step

Once you have confirmed that feature matches look sensible, return to
`biber()` for corpus-level counts or move on to the multidimensional analysis
workflows shown in the other vignettes.